---
title: "BIO 620 Week 6 – Assignment 2"
author: "Haille Bright"
date: "2026-09-29"
output: word_document
---
## Assignment 2
Estimate the proportion of individuals who like chocolate using Bayesian statistics.
```{r}
# Number of people surveyed
total <- 50

# Number of people who like chocolate
successes <- 45
```
```{r}
# Set the parameters for the prior beta distribution
a <- 52.22
b <- 9.52

# Plot the prior distribution
p <- seq(0, 1, length = 1000)
plot(p, dbeta(p, a, b),
     type = "l",
     xlab = "Proportion who like chocolate",
     ylab = "Density",
     main = "Prior Distribution")
```
```{r}
# Calculate the likelihood
likelihood <- dbinom(successes, size = total, prob = p)

# Plot the likelihood
plot(p, likelihood,
     type = "l",
     xlab = "Proportion who like chocolate",
     ylab = "Likelihood",
     main = "Likelihood Distribution")
```
```{r}
# Calculate posterior parameters
posterior_a <- a + successes
posterior_b <- b + (total - successes)

# Plot the posterior distribution
plot(p, dbeta(p, posterior_a, posterior_b),
     type = "l",
     xlab = "Proportion who like chocolate",
     ylab = "Density",
     main = "Posterior Distribution")
```
```{r}
# Calculate the posterior mean
posterior_mean <- posterior_a / (posterior_a + posterior_b)

posterior_mean
```
The survey found that 45 out of 50 individuals liked chocolate, giving an observed proportion of 0.90. The prior distribution estimated that the proportion was most likely around 0.85. After combining the prior information with the survey data, the posterior mean was approximately 0.87. Based on the Bayesian analysis, the estimated proportion of individuals who like chocolate is about 87%.