---
title: "BIO620 Week 6 – Assignment 3"
author: "Haille Bright"
date: "2026-09-29"
output: word_document
---

## Assignment 3
Bayesian analysis of unknown dice rolls.
```{r}
# Possible dice types
dice <- c(4, 6, 8, 10, 12, 20)

# Dice rolls for each player
benji_rolls <- c(8, 8, 5, 16, 6, 8)
marcus_rolls <- c(12, 8, 2, 1, 4, 7)
carrie_rolls <- c(5, 8, 7, 3, 2, 6)
lucy_rolls <- c(3, 5, 2, 3, 4, 3)
```
```{r}
# Prior probabilities based on each player's dice collection
benji_prior <- c(25, 5, 10, 15, 20, 25) / 100
marcus_prior <- c(4, 8, 16, 32, 25, 15) / 100
carrie_prior <- c(20, 19, 20, 20, 1, 20) / 100
lucy_prior <- c(4, 6, 8, 10, 12, 60) / 100
```
```{r}
# Function to calculate log-likelihood for a die
dice_loglik <- function(rolls, sides) {
  if (any(rolls > sides)) {
    return(-Inf)
  } else {
    return(length(rolls) * log(1 / sides))
  }
}

# Calculate Benji's log-likelihood for each die type
benji_loglik <- sapply(dice, function(x) dice_loglik(benji_rolls, x))

# Display results
data.frame(
  Die = paste0("d", dice),
  LogLikelihood = benji_loglik
)
```
```{r}
# Convert log-likelihoods to likelihoods
benji_likelihood <- exp(benji_loglik)

# Combine likelihood with Benji's prior
benji_posterior <- benji_likelihood * benji_prior

# Normalize posterior probabilities
benji_posterior <- benji_posterior / sum(benji_posterior)

# Display posterior probabilities
data.frame(
  Die = paste0("d", dice),
  Posterior = benji_posterior
)
```
```{r}
# Prepare log-likelihood values for plotting
benji_plot_loglik <- benji_loglik
benji_plot_loglik[is.infinite(benji_plot_loglik)] <- 
  min(benji_loglik[is.finite(benji_loglik)]) - 5

# Plot Benji's log-likelihood
barplot(benji_plot_loglik,
        names.arg = paste0("d", dice),
        xlab = "Die Type",
        ylab = "Log-Likelihood",
        main = "Benji: Log-Likelihood")
```
```{r}
# Plot Benji's posterior probabilities
barplot(benji_posterior,
        names.arg = paste0("d", dice),
        xlab = "Die Type",
        ylab = "Posterior Probability",
        main = "Benji: Posterior Probability",
        ylim = c(0, 1))
```
```{r}
# Calculate Marcus's log-likelihood for each die type
marcus_loglik <- sapply(dice, function(x) dice_loglik(marcus_rolls, x))

# Display results
data.frame(
  Die = paste0("d", dice),
  LogLikelihood = marcus_loglik
)
```
```{r}
# Convert log-likelihoods to likelihoods
marcus_likelihood <- exp(marcus_loglik)

# Combine likelihood with Marcus's prior
marcus_posterior <- marcus_likelihood * marcus_prior

# Normalize posterior probabilities
marcus_posterior <- marcus_posterior / sum(marcus_posterior)

# Display posterior probabilities
data.frame(
  Die = paste0("d", dice),
  Posterior = marcus_posterior
)
```
```{r}
# Prepare log-likelihood values for plotting
marcus_plot_loglik <- marcus_loglik
marcus_plot_loglik[is.infinite(marcus_plot_loglik)] <-
  min(marcus_loglik[is.finite(marcus_loglik)]) - 5

# Plot Marcus's log-likelihood
barplot(marcus_plot_loglik,
        names.arg = paste0("d", dice),
        xlab = "Die Type",
        ylab = "Log-Likelihood",
        main = "Marcus: Log-Likelihood")
```
```{r}
# Plot Marcus's posterior probabilities
barplot(marcus_posterior,
        names.arg = paste0("d", dice),
        xlab = "Die Type",
        ylab = "Posterior Probability",
        main = "Marcus: Posterior Probability",
        ylim = c(0, 1))
```
```{r}
# Calculate Carrie's log-likelihood for each die type
carrie_loglik <- sapply(dice, function(x) dice_loglik(carrie_rolls, x))

# Display results
data.frame(
  Die = paste0("d", dice),
  LogLikelihood = carrie_loglik
)
```
```{r}
# Convert log-likelihoods to likelihoods
carrie_likelihood <- exp(carrie_loglik)

# Combine likelihood with Carrie's prior
carrie_posterior <- carrie_likelihood * carrie_prior

# Normalize posterior probabilities
carrie_posterior <- carrie_posterior / sum(carrie_posterior)

# Display posterior probabilities
data.frame(
  Die = paste0("d", dice),
  Posterior = carrie_posterior
)
```
```{r}
# Prepare log-likelihood values for plotting
carrie_plot_loglik <- carrie_loglik
carrie_plot_loglik[is.infinite(carrie_plot_loglik)] <-
  min(carrie_loglik[is.finite(carrie_loglik)]) - 5

# Plot Carrie's log-likelihood
barplot(carrie_plot_loglik,
        names.arg = paste0("d", dice),
        xlab = "Die Type",
        ylab = "Log-Likelihood",
        main = "Carrie: Log-Likelihood")
```
```{r}
# Plot Carrie's posterior probabilities
barplot(carrie_posterior,
        names.arg = paste0("d", dice),
        xlab = "Die Type",
        ylab = "Posterior Probability",
        main = "Carrie: Posterior Probability",
        ylim = c(0, 1))
```
```{r}
# Calculate Lucy's log-likelihood for each die type
lucy_loglik <- sapply(dice, function(x) dice_loglik(lucy_rolls, x))

# Display results
data.frame(
  Die = paste0("d", dice),
  LogLikelihood = lucy_loglik
)
```
```{r}
# Convert log-likelihoods to likelihoods
lucy_likelihood <- exp(lucy_loglik)

# Combine likelihood with Lucy's prior
lucy_posterior <- lucy_likelihood * lucy_prior

# Normalize posterior probabilities
lucy_posterior <- lucy_posterior / sum(lucy_posterior)

# Display posterior probabilities
data.frame(
  Die = paste0("d", dice),
  Posterior = lucy_posterior
)
```
```{r}
# Prepare log-likelihood values for plotting
lucy_plot_loglik <- lucy_loglik
lucy_plot_loglik[is.infinite(lucy_plot_loglik)] <-
  min(lucy_loglik[is.finite(lucy_loglik)]) - 5

# Plot Lucy's log-likelihood
barplot(lucy_plot_loglik,
        names.arg = paste0("d", dice),
        xlab = "Die Type",
        ylab = "Log-Likelihood",
        main = "Lucy: Log-Likelihood")
```
```{r}
# Plot Lucy's posterior probabilities
barplot(lucy_posterior,
        names.arg = paste0("d", dice),
        xlab = "Die Type",
        ylab = "Posterior Probability",
        main = "Lucy: Posterior Probability",
        ylim = c(0, 1))
```
## Results and Interpretation

For Benji, the d20 had a posterior probability of 1.00 because his roll of 16 ruled out all of the smaller dice.

For Marcus, the d12 had the highest posterior probability at approximately 0.973, while the d20 had a posterior probability of approximately 0.027. His roll of 12 ruled out the d4, d6, d8, and d10.

For Carrie, the d8 had the highest posterior probability at approximately 0.787, followed by the d10 at approximately 0.206. The d12 and d20 had very low posterior probabilities.

For Lucy, the d6 had the highest posterior probability at approximately 0.739, followed by the d8 at approximately 0.175. Although Lucy's prior probability was highest for the d20 because 60 of her 100 dice were d20s, the observed rolls provided stronger evidence for a smaller die.

Overall, the Bayesian analysis combined the likelihood of the observed rolls with the prior probability based on the number of each die type in each player's collection.
