---
title: "BIO 620 Week 5 Assignments 2 and 3"
author: "Haille Bright"
date: "2026-09-25"
output: word_document
---
## Assignment 2: Argiope Fertilization Dataset
```{r}
argiope <- read.csv("Argiope_fertilization_small.csv")
```

```{r}
str(argiope)
```
```{r}
summary(argiope)
```
```{r}
model1 <- lm(NumFertilized ~ InsertionDuration + MaleSize + FemaleSize + NumEggs + NumSacs,
             data = argiope)

summary(model1)
```
```{r}
plot(argiope$InsertionDuration, argiope$NumFertilized,
     xlab = "Insertion Duration",
     ylab = "Number of Fertilized Eggs",
     main = "Insertion Duration vs. Number of Fertilized Eggs")

abline(lm(NumFertilized ~ InsertionDuration, data = argiope))
```
```{r}
par(mfrow = c(2, 2))
plot(model1)
par(mfrow = c(1, 1))
```
The multivariate linear model showed that insertion duration was significantly associated with the number of fertilized eggs (p = 0.008). The relationship was negative, meaning that longer insertion durations were associated with fewer fertilized eggs when male size, female size, number of eggs, and number of egg sacs were accounted for. The number of eggs was also a significant predictor (p < 0.001), while male size, female size, and number of egg sacs were not significant predictors. The overall model was significant (p < 0.001) and explained approximately 71% of the variation in the number of fertilized eggs.


## Assignment 3: Quine Dataset
```{r}
quine <- read.csv("quine.csv")

quine$Eth <- as.factor(quine$Eth)
quine$Sex <- as.factor(quine$Sex)
quine$Age <- as.factor(quine$Age)
quine$Lrn <- as.factor(quine$Lrn)

str(quine)
```
```{r}
quine$Eth <- as.factor(quine$Eth)
quine$Sex <- as.factor(quine$Sex)
quine$Age <- as.factor(quine$Age)
quine$Lrn <- as.factor(quine$Lrn)

str(quine)
```
```{r}
quine_model <- aov(Days ~ Eth + Sex + Age + Lrn, data = quine)

summary(quine_model)
```
```{r}
TukeyHSD(quine_model)
```
```{r}
boxplot(Days ~ Eth, data = quine,
        xlab = "Ethnicity",
        ylab = "Days Absent",
        main = "Days Absent by Ethnicity")

boxplot(Days ~ Age, data = quine,
        xlab = "Age Group",
        ylab = "Days Absent",
        main = "Days Absent by Age Group")
```
The ANOVA showed that ethnicity and age were significant predictors of the number of days absent. Ethnicity was significant (p < 0.001), while age was also significant (p = 0.031). Sex (p = 0.275) and learner status (p = 0.087) were not significant predictors. The Tukey HSD test showed that the N ethnicity group had significantly fewer days absent than the A group (p < 0.001). For age, the only significant difference was between the F1 and F2 groups (p = 0.045), with the F2 group having more days absent than the F1 group.
