---
title: "Week 5 Assign. 3"
author: "Haille Bright"
date: "2026-09-25"
output: word_document
---
## Assignment 3: Quine Dataset
```{r}
quine <- read.csv("quine.csv")

quine$Eth <- as.factor(quine$Eth)
quine$Sex <- as.factor(quine$Sex)
quine$Age <- as.factor(quine$Age)
quine$Lrn <- as.factor(quine$Lrn)

str(quine)
```
```{r}
quine$Eth <- as.factor(quine$Eth)
quine$Sex <- as.factor(quine$Sex)
quine$Age <- as.factor(quine$Age)
quine$Lrn <- as.factor(quine$Lrn)

str(quine)
```
```{r}
quine_model <- aov(Days ~ Eth + Sex + Age + Lrn, data = quine)

summary(quine_model)
```
```{r}
TukeyHSD(quine_model)
```
```{r}
boxplot(Days ~ Eth, data = quine,
        xlab = "Ethnicity",
        ylab = "Days Absent",
        main = "Days Absent by Ethnicity")

boxplot(Days ~ Age, data = quine,
        xlab = "Age Group",
        ylab = "Days Absent",
        main = "Days Absent by Age Group")
```
The ANOVA showed that ethnicity and age were significant predictors of the number of days absent. Ethnicity was significant (p < 0.001), while age was also significant (p = 0.031). Sex (p = 0.275) and learner status (p = 0.087) were not significant predictors. The Tukey HSD test showed that the N ethnicity group had significantly fewer days absent than the A group (p < 0.001). For age, the only significant difference was between the F1 and F2 groups (p = 0.045), with the F2 group having more days absent than the F1 group.
