---
title: "Week7Assignment4"
author: "Haille Bright"
date: "2026-10-06"
output: word_document
---
## Assignment 4: Mole Rat Dataset

```{r}
molerat <- read.csv(file.choose())

head(molerat)
str(molerat)
summary(molerat)
```
```{r}
molerat$caste <- as.factor(molerat$caste)

str(molerat)
```

```{r}
# Gamma GLM models

m1_m <- glm(lnenergy ~ lnmass,
            family = Gamma(link = "log"), data = molerat)

m2_m <- glm(lnenergy ~ caste,
            family = Gamma(link = "log"), data = molerat)

m3_m <- glm(lnenergy ~ lnmass + caste,
            family = Gamma(link = "log"), data = molerat)

m4_m <- glm(lnenergy ~ lnmass * caste,
            family = Gamma(link = "log"), data = molerat)
```

```{r}
AIC(m1_m, m2_m, m3_m, m4_m)
```

```{r}
aic_molerat <- AIC(m1_m, m2_m, m3_m, m4_m)

aic_molerat$Delta_AIC <- aic_molerat$AIC - min(aic_molerat$AIC)

aic_molerat
```
```{r}
summary(m3_m)
```
```{r}
anova(m3_m, test = "Chisq")
```
Model selection showed that the best Gamma GLM included lnmass and caste as predictors of lnenergy (AIC = 18.79). Both lnmass (p < 0.001) and caste (p = 0.008) had significant effects on lnenergy. The positive effects suggest to me that lnenergy increased with lnmass and was higher for workers compared with the lazy caste. The interaction model had a similar AIC, but the additive model was selected because it had the lowest AIC and was simpler.
