---
title: 'Week 5 Assignment #2'
author: "Haille Bright"
date: "2026-09-25"
output: word_document
---
## Assignment 2: Argiope Fertilization Dataset
```{r}
str(argiope)
```
```{r}
summary(argiope)
```
```{r}
model1 <- lm(NumFertilized ~ InsertionDuration + MaleSize + FemaleSize + NumEggs + NumSacs,
             data = argiope)

summary(model1)
```
```{r}
plot(argiope$InsertionDuration, argiope$NumFertilized,
     xlab = "Insertion Duration",
     ylab = "Number of Fertilized Eggs",
     main = "Insertion Duration vs. Number of Fertilized Eggs")

abline(lm(NumFertilized ~ InsertionDuration, data = argiope))
```
```{r}
par(mfrow = c(2, 2))
plot(model1)
par(mfrow = c(1, 1))
```
The multivariate linear model showed that insertion duration was significantly associated with the number of fertilized eggs (p = 0.008). The relationship was negative, meaning that longer insertion durations were associated with fewer fertilized eggs when male size, female size, number of eggs, and number of egg sacs were accounted for. The number of eggs was also a significant predictor (p < 0.001), while male size, female size, and number of egg sacs were not significant predictors. The overall model was significant (p < 0.001) and explained approximately 71% of the variation in the number of fertilized eggs.

