---
title: "Week 7 Assignment 2"
author: "Haille Bright"
date: "2026-10-06"
output: word_document
---

## Assignment 2: Species Dataset

```{r}
species <- read.csv(file.choose())
```
```{r}
head(species)
```
```{r}
str(species)
summary(species)
```
```{r}
species$pH <- as.factor(species$pH)

str(species)
```
```{r}
# Poisson GLM models

m1 <- glm(Species ~ Biomass, 
          family = poisson, data = species)

m2 <- glm(Species ~ pH, 
          family = poisson, data = species)

m3 <- glm(Species ~ Biomass + pH, 
          family = poisson, data = species)

m4 <- glm(Species ~ Biomass * pH, 
          family = poisson, data = species)
```
```{r}
AIC(m1, m2, m3, m4)
```
```{r}
aic_values <- AIC(m1, m2, m3, m4)

aic_values$Delta_AIC <- aic_values$AIC - min(aic_values$AIC)

aic_values
```
```{r}
summary(m4)
```
```{r}
anova(m4, test = "Chisq")
```
```{r}
deviance(m4) / df.residual(m4)
```
The model selection showed that the best model was the interaction model including biomass, pH, and biomass × pH (AIC = 514.39). Both biomass and pH had significant effects on species count, and there was also a significant interaction between biomass and pH (p < 0.001). This suggests that the relationship between biomass and species count changes depending on pH level. The dispersion value was 0.99, indicating that overdispersion was not a concern for the Poisson model.
